Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-22 3dimpvt/paper.tex
completed v2 changes

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and choosing an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
intersect them miniimally and are likely to be less noticeable.

@@

These two criteria can be met by maximizing the function

\[
\frac{1}{d} \cdot (-1 \cdot C) \cdot N
\]

which minimizes camera angle $\alpha$ and distance $d$ as shown in
Figure \ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{Images are selected by minimizing camera angle $\alpha$ and
    distance $d$.}
  \label{fig:scoringFunction}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select different
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If a tile is not visible in
images chosen by its neighbors, using similar images is likely to
result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function in Figure
\ref{fig:scoringFunction}. Rather than blindly reusing this image, we
ensure it meets a threshold, which we set to $\alpha < 45^\circ$, to
be considered a good image.

If no good image is found, we then check our second subset of images,
which consists of images that were taken near the images in the first
subset, both spatially and temporally. These images are not the same
as the ones used for neighboring tiles, but they were taken at a
similar location and time, suggesting that their localization and
projection are very similar. Again, if no good image is found
according to the same threshold, we then must search the entire set of
candidate images.

@@

The result of this caching approach is shown in figure
\ref{fig:caching}. As compared to \ref{fig:directMapping},
discontinuities have been reduced overall, but the amount of remaining
seams suggests that image selection alone cannot produce seamless
textures. Camera matrices, or the image projections themselves have to
be adjusted in order to reliably generate clean textures.

@@

Rather than perform camera or image adjustments across the many
thousands of images acquired in a typical data collection, we opt to
work with the more limited set of images corresponding to those chosen
by the direct mapping approach, without caching. This set of images
constitutes a good candidate set for generating a final seamless
texture since it meets three important criteria. First, it contains at least one image that covers each tile on our
plane; this ensures no holes in our final texture. Second, since
images are all selected according to the same scoring function in
Figure \ref{fig:scoringFunction}, they are taken at as much of a head-on angle as possible and should project
onto the plane in similar ways. Third, as a side result of the scoring
function, selected images are only good candidates for the tiles near
their center of projection. Thus, there should be plenty of overlap
between selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards refining and
combining their projections.

@@

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}
\begin{figure}
  \centering
  \includegraphics[width=3in]{Graph_crop.pdf}
  \caption{The graph-based localization refinement algorithm from [11]
    suffers from the problem of compounding errors. }
  \label{fig:graph}
\end{figure}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack
localization\cite{liu2010indoor,chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem shown in
Figure \ref{fig:graph}.

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
Our proposed approach towards texture mapping is a two-step
process. First, with the same input set of images as described in
Section \ref{sec:existingApproaches}, we perform image rotation and
shifting in order to maximize SIFT feature matches between images. We
then project these images onto our plane, applying textures either
with the tile caching method from Section
\ref{sec:mappingWithCaching}, or with the more specialized method
described in Section \ref{sec:seamMinimization}.

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and
match each other with far greater accuracy. Applying the tile caching method from Section \ref{sec:mappingWithCaching} on these re-localized images, results in the significant improvements shown in Figure \ref{fig:shifted}.

@@

As mentioned earlier, our backpack system has 2 cameras with
$180^\circ$ fisheye lenses facing to the right and left. Since the
backpack operator only takes care to fully scan each wall and not
necessarily the entirety of each ceiling and floor, cameras at higher
angles with respect to plane normal vectors must be used for texturing
large areas of floor and ceiling planes. These high camera angles
translate into images that span extremely large areas once projected
onto their planes. Such image projections have good scores (from
Figure \ref{fig:scoringFunction}) for certain areas on the plane, but
much worse scores for other areas. This can be seen in Figure
\ref{fig:highCameraAngle}. The tiling approach used in Section
\ref{sec:simpleTextureMapping} was thus a good choice for texturing,
as it allowed us to only use the parts of image projections that were
good choices for their respective plane locations.

@@

\subsection{Occlusion Masking}
\label{sec:occlusionMasking}
To begin with, we need to ensure that our images contain only content
that should be mapped onto the target plane in question. The tiling
approach used previously only checks occlusion for each tile as it is
being textured. For our new approach, we use entire images, so we need
to perform occlusion checks over the entirety of each image to
determine available areas for texture mapping.

@@

With the cost function defined, we must now select the set of images
for which the overall cost function is minimized. Since nearly all the
images for our optimal case are head on, the best strategy to minimize
seams is to choose as few images as possible while texturing a given
plane. Thus,to cover the entirety of a plane, our problem can be
defined as minimally covering a polygon i.e. the plane, using other
polygons of arbitrary geometry i.e. our image projections, with the
added constraint of minimizing our cost function between chosen
images.  This is a complex problem, though we can take a number of
steps to simplify it. Given that our wall-texture candidate images are
all taken from a head-on angle, and assuming only minor rotations are
made during localization refinement, we can reason that their
projections onto the plane are approximately rectangular, as in Figure
\ref{sec:goodCameraAngle}. By discarding the minor excess texture and
cropping them all to be rectangular, our problem becomes the
conceptually simpler one of filling a polygon with rectangles, such
that the sum of all edge costs between each pair of rectangles is
minimal. We thus also retain the advantages of working with
rectangular units, as explained in section
\ref{sec:simpleTextureMapping}.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by a chosen image, we are left with a hole where no
image is chosen to texture.

SOMETHING HERE ABOUT CRAP CRAP

Rather than reverting to a 2D-coverage problem, we can elect to simply
fill the hole by selecting images to fill it in a greedy fashion with
respect to edge costs of the same cost function.

@@

In the context of the backpack modeling system, we apply the seam
minimization approach on walls, due to its superior output when
provided with head-on images. Floors and ceilings however, given their
images taken at oblique angles, as shown in Figure \ref{fig:floor_suboptimal}, are textured using the tile caching method.

@@

We apply alpha blending to blend pixels
linearly across overlapping regions. Figure
\ref{fig:pipeline} shows our entire texture mapping pipeline and demonstrates the
final blended output of both approaches.

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our final texture processing pipeline, with the final
    output of both approaches.}
  \label{fig:pipeline}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{4thfloor21.jpg}
  \includegraphics[width=3in]{4thfloor61.jpg}
  \includegraphics[width=2in]{4thfloor8.jpg}
  \includegraphics[width=3in]{fullmodel.png}
  \caption{Examples of our final texture mapping output for walls,
    floors, and ceilings}
  \label{fig:results}
\end{figure}

>>>>

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
generally about 5 pixels across, and choosing an image to texture each
tile with. We choose to work with rectangular units to ensure that
borders between any two distinct images in our final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

@@

These two criteria can be met by maximizing the function shown in
Figure \ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{Images with minimal camera angle $\alpha$ and distance $d$
    are selected by maximizing the function}
  \begin{center}$\frac{1}{d} \cdot (-1 \cdot C) \cdot N$ \end{center}
  \label{fig:scoringFunction}
\end{figure}

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select different
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If this is not possible
because a tile is not visible in images chosen by its neighbors, using
similar images is likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function in
\ref{fig:scoringFunction}. Rather than blindly reusing this image, we
ensure it meets a threshold, which we set to $\alpha < 45^\circ$, to
be considered a good image.

If no satisfactory image is found, we then check our second subset of
images, which consists of images that were taken near the images in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but they were taken
at a similar location and time, suggesting that their localization and
projection are very similar. Again, if no good image is found
according to the same threshold, we must then search the entire set of
candidate images.

@@

The result of this caching approach is shown in Figure
\ref{fig:caching}. As compared to Figure \ref{fig:directMapping},
discontinuities have been reduced overall, but the amount of remaining
seams suggests that image selection alone cannot produce seamless
textures. Camera matrices, or the image projections themselves have to
be adjusted in order to reliably generate clean textures.

@@

Rather than perform camera or image adjustments across the many
thousands of images acquired in a typical data collection, we opt to
work with the more limited set of images corresponding to those chosen
by the direct mapping approach, without caching. This set of images
constitutes a good candidate set for generating a final seamless
texture since it meets three important criteria. First, each tile on our planeis covered by at least one image in this set; this ensures no
holes in our final texture. Second, since images are all selected
according to the same scoring function in Figure
\ref{fig:scoringFunction}, they are taken at as much of a head-on
angle as possible and should project onto the plane in similar
ways. Third, as a side result of the scoring function, selected images
are only good candidates for the tiles near their center of
projection. Thus, there should be plenty of overlap between selected
images, allowing for some degree of shifting without resulting in
holes, as well as area for blending between them. With this set of
images, we now review two existing approaches towards refining and
combining their projections.

@@

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}
\begin{figure}
  \centering
  \includegraphics[width=3in]{Graph_crop.pdf}
  \caption{The graph-based localization refinement algorithm from \cite{chen2010indoor}
    suffers from the problem of compounding errors. }
  \label{fig:graph}
\end{figure}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack
localization\cite{liu2010indoor, chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem as shown in
Figure \ref{fig:graph}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}
\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach contains 4 steps, as seen in
\ref{fig:pipeline}.  First, we obtain a set of images to work with, as
already described in Section \ref{sec:existingApproaches}. Next, we perform
image rotation and shifting in order to maximize SIFT feature matches
in overlapping areas between images. We then project these images onto
our plane, applying textures either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method described in Section \ref{sec:seamMinimization}. We then finish
by applying linear alphar blending to smooth out any remaining seams.

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these re-localized images
results in the significant improvements shown in Figure
\ref{fig:shifted}, as compared to Figure \ref{fig:caching}.

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{highCameraAngle.pdf}
  \caption{The image on the left projects head-on not only onto point A, but most visible areas of the wall that point A is on. The image's projection onto point B will be at a high angle, but it will still be much more usable than the projection onto point C.}
  \label{fig:highCameraAngle}
\end{figure}

As mentioned earlier, our backpack system has 2 cameras with
$180^\circ$ fisheye lenses facing to the right and left. Since the
backpack operator only takes care to fully scan each wall and not
necessarily the entirety of each ceiling and floor, cameras at higher
angles with respect to plane normal vectors must be used for texturing
large areas of floor and ceiling planes. These high camera angles
translate into images that span extremely large areas once
projected. Such image projections have good scores (from Figure
\ref{fig:scoringFunction}) for certain areas on the plane, but much
worse scores for other areas. This can be seen in Figure
\ref{fig:highCameraAngle}. The tiling approach used in Section
\ref{sec:simpleTextureMapping} was thus a good choice for texturing,
as it allowed us to only use the parts of image projections that were
good choices for their respective plane locations.

@@

\subsection{Occlusion Masking}
\label{sec:occlusionMasking}
To begin with, we need to ensure that our images contain only content
that should be mapped onto the target plane in question. The tiling
approach used previously only checks occlusion for each tile as it is
being textured. For our new approach, we prefer entire images, so we need
to perform occlusion checks over the entirety of each image to
determine available areas for texture mapping.

@@

With the cost function defined, we must now select the set of images
for which the overall cost function is minimized. Since nearly all the
images for our optimal case are head on, the best strategy to minimize
seams is to choose as few images as possible while texturing a given
plane. Thus,to cover the entirety of a plane, our problem can be
defined as minimally covering a polygon i.e. the plane, using other
polygons of arbitrary geometry i.e. our image projections, with the
added constraint of minimizing our cost function between chosen
images.  This is a complex problem, though we can take a number of
steps to simplify it. Given that our wall-texture candidate images are
all taken from a head-on angle, and assuming only minor rotations are
made during localization refinement, we can reason that their
projections onto the plane are approximately rectangular, as in Figure
\ref{fig:highCameraAngle}. By discarding the minor excess texture and
cropping them all to be rectangular, our problem becomes the
conceptually simpler one of filling a polygon with rectangles, such
that the sum of all edge costs between each pair of rectangles is
minimal. We thus also retain the advantages of working with
rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes where no
images are selected to texture. Rather than reverting to a 2D-coverage problem, we simply repeat the process, but first modify the edges chosen by our shortest path solution such that they have higher cost than all other edges in our DAG. This ensures that these edges will not be chosen again unless they are the only available option. After solving the shortest path problem with these modified edge costs, we receive a new set of chosen images, and only retain those that cover areas not covered by the first set. This process can be repeated as many times as needed, until holes are no longer present. This method is not as optimal as a 2D-coverage solution would be, but it is quick to run and adequately handles the few holes we run into.

@@

In the context of the backpack modeling system, we apply the seam
minimization approach on walls, due to its superior output when
provided with head-on images. Floors and ceilings however, given their
many images taken at oblique angles, as shown in Figure
\ref{fig:highCameraAngle}, are textured using the tile caching
method.

@@

After blending pixels linearly across overlapping
regions using alpha blending, our texture mapping is complete. Figure \ref{fig:pipelineimages} shows each incremental step of our texture mapping approach, as well as the final blended output resulting from both of our methods.

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipelineimages.pdf}
  \caption{Textures corresponding to each step in our texture mapping pipeline.}
  \label{fig:pipelineimages}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{4thfloor21.jpg}
  \includegraphics[width=3in]{4thfloor61.jpg}
  \includegraphics[width=3in]{wall3final.jpg}
  \includegraphics[width=2in]{4thfloor8.jpg}
  \includegraphics[width=3in]{floorfinal.jpg}
  \includegraphics[width=3in]{fullmodel.png}
  \caption{Examples of our final texture mapping output for walls,
    floors, and ceilings}
  \label{fig:results}
\end{figure}
